---
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/grout)](https://CRAN.R-project.org/package=grout)
[![R-CMD-check](https://github.com/hypertidy/grout/workflows/R-CMD-check/badge.svg)](https://github.com/hypertidy/grout/actions)
<!-- badges: end -->
  
# grout

Abstract tiling schemes. 


Given a grid, impose a tiling scheme. The dangle is overlap the tiles impose (in pixels). 

We use the term "block" to refer to the size of each tile in pixels. 

Consider a raster 16 x 12, with 4 x 4 tiling there is no overlap. 

```{r overlap}
library(grout)
r0 <- raster::raster(matrix(0, 12, 16), xmn = 0, xmx = 16, ymn = 0, ymx = 12)
tiles(r0, blockX = 4, blockY = 4)

```

But, if our raster has an dimension that doesn't divide neatly into the block size, 
then there is some tile overlap. This should work for any raster dimension and any arbitrary tile size. 


```{r dangle}
r1 <- raster::raster(matrix(0, 13, 15), xmn = 0, xmx = 15, ymn = 0, ymx = 13)
tiles(r1, 4, 4)

```


```{r grid}

r <- raster::raster(matrix(1:12, 30, 44), xmn= 0,xmx = 4, ymn = 0, ymx = 3)


(t1 <- tiles(r, blockX = 12, blockY = 12))
plot(t1)

(t2 <- tiles(volcano, 12, 16))

plot(t2)
```

We can generate a table of offset indexes, for use in reading from GDAL (say). 

```{r offsets}
tile_index(t2)
```

We can generate polygons from these. 

```{r poly}
p <- as_polys(t1)

sp::plot(p)
```

Or just plot the scheme. 

```{r plot}
plot(t1)
```

## What for? 

This gives us fine control over the exact nature of the data we can read from large sources. 

Consider this large image online: 

```{r large}
u <- "/vsicurl/https://s3-us-west-2.amazonaws.com/planet-disaster-data/hurricane-harvey/SkySat_Freeport_s03_20170831T162740Z3.tif"

im <- raster::brick(u)
im
```

The raster reader gives us a useful brick object that can be operated with, if we crop it only those 
cell values are read - but we have no idea about the underlying tiling of the data source itself. 

With GDAL more directly we can find the underlying tile structure, which tells us about the 256x256 tiling
scheme. 

```{r tile}
info <- vapour::vapour_raster_info(u)

info["tilesXY"]
```

Now with grout we can actually generate the tile scheme and work with it, let's say we know that we want a region near tile number 6500. Using the raster version of the tiles we can find the adjacent tile cell numbers, then use that to crop the original source. 

```{r scheme}
(tile0 <- tiles(im, info$tilesXY[1], info$tilesXY[2]))

polys <- as_polys(tile0)
cells <- raster::adjacent(tile0$tileraster, 6500, include = TRUE, directions = 8)[, "to"]
rr <- raster::crop(im, polys[cells, ])
```

Finally, we have an in-memory raster of the original source data for very specific tiles. 

```{r specific}
rr

raster::plotRGB(rr)
sp::plot(polys[cells, ], add = TRUE, border = "firebrick")
text(sp::coordinates(polys[cells, ]), lab = cells, col = "yellow")
```

## TODO

* remove need for using sp polygons
* set tools for cropping that use the index, not spatial extent (i.e. extent(x, x0, x1, y0, y1))
* remove use of sp and raster internally for the data structures, just store the information about the grid/s


---


## Code of Conduct
  
Please note that the grout project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
